### 2016-06-11 fig_layers.tex
Computing Stack layers diagram is added

\begin{figure}
\centering
\begin{tikzpicture}[scale=0.50,node distance=3pt,outer sep=0pt,
platformlayer/.style={
  draw=darkgray,
  fill=nodecolor,
  rounded corners,
  text width=4.5cm,
  font={\sffamily\bfseries\color{black}},
  align=center,
  text height=9pt,
  text depth=6pt},
]
\node[platformlayer] (APP) {Application};
\node[platformlayer,below=of APP](OS) {Operating System};
\node[platformlayer,below=of OS] (HYP) {Hypervisor};
\node[platformlayer,below=of HYP](BUS){BUS and Interconnect};
\node[platformlayer,below=of BUS](FW) {Firmware};
\node[platformlayer,below=of FW](HW) {Hardware};

\begin{pgfonlayer}{background}
\draw[platformlayer,draw=black,fill=nodecolor!40]([xshift=-\framesep,yshift=1\framesep]current bounding box.north west)
rectangle ([xshift=\framesep,yshift=-\framesep]current bounding box.south east) 
node[below] at (0,2) {};

\end{pgfonlayer}
\end{tikzpicture}
\caption{Computing Platform Layers}
\label{fig:layers}
\end{figure}